Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.tex
copy editing towards the end

\caption[Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green)]{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green). The projection problem is a 2D projection with three variables using the holes index. The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus. The full video of the animation can be found in the html version of the paper.}\label{fig:toy-torus}
\end{figure}
\end{Schunk}

@@

Random variables with different distributions have been simulated and
the distributional form of each variable is presented in equations
\ref{eq:sim-norm} to \ref{eq:sim-x7}. Variables \texttt{x1}, \texttt{x8}
to \texttt{x10} are normal noise with zero mean and unit variance and
\texttt{x2} to \texttt{x7} are normal mixtures with varied weights and
locations. All the variables have been scaled to have an overall unit
variance before projection pursuit. The holes index
\citep{cook2008grand}, used for detecting bimodality of the variables,
is used throughout the examples unless otherwise specified.

@@

An example of non-monotonic interpolation has been given in Figure
\ref{fig:toy-interp}: a path that passes bases with a higher index value
than the target one. For SA, a non-monotonic interpolation is justified
since target bases do not necessarily have a higher index value than the
current one, while this is not the case for CRS. The original trace plot
for a 2D projection problem, optimised by CRS, is shown on the left
panel of Figure \ref{fig:interruption} and one can observe that the
non-monotonic interpolation has undermined the optimiser to realise its
full potential. Hence, an interruption is implemented to stop at the
best basis found in the interpolation. The right panel of Figure
\ref{fig:interruption} shows the trace plot after implementing the
interruption and while the first two interpolations are identical, the
basis at time 61 has a higher index value than the target in the third
interpolation. Rather than starting the next iteration from the target
basis on time 65, CRS starts the next iteration at time 61 on the right
panel and reaches a better final basis.

@@

\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS. On the left panel, basis with a higher index value is found during the interpolation but not used. On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel.}\label{fig:interruption}
\end{figure}
\end{Schunk}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  the distance between the basis found and the current needs to be
  larger than 1e-3;
\item
  the relative change of the index value needs to be larger than 1e-5;
  and
\item
  the alpha parameter needs to be larger than 0.01
\end{enumerate}

Figure \ref{fig:polish} presents the projected data and trace plot of a
2D projection, optimised by CRS and followed by the polish step. The top
row shows the initial projection, the final projection after CRS, and
the final projection after polish, respectively The end basis found by
CRS reveals the four clusters in the data, but the edges of each cluster
are not clean-cut. Polish works with this end basis and further pushes
the index value to produce clearer edges of the cluster, especially
along the vertical axis.

@@

\caption{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index. The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value. The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis. Note that the parameter \code{max.tries} is set to 400 in this experiment for CRS to do its best. }\label{fig:polish}
\end{figure}
\end{Schunk}

@@

The holes index function used for all the examples before this section
produces a smooth interpolation, while this is not the case for all the
indexes. An example of a noisy index function for 1D projections
compares the projected data, \(\mathbf{Y}_{n \times 1}\), to a randomly
generated normal distribution, \(\mathcal{N}_{n \times 1}\), using the
Kolmogorov test. Let \(F_{.}(n)\) be the ECDF function (empirical
cummulative distribution function), with two possible subscripts \(Y\)
and \(\mathcal{N}\) representing the projected and randomly generated
data, and \(n\) denoting the number of observation, the Kolmogorov index
\(I^{nk}(n)\), is defined as:

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  whether any optimiser fails to optimise this non-smooth index; and
\item
  whether the optimisers can find the global optimum despite the
  presence of a local optimum
\end{enumerate}

Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of
all three optimisers and as expected, none of the interpolated paths are
smooth. There is barely any improvement made by PD, indicating its
failure in optimising non-smooth index functions. While CRS and SA have
managed to get close to the index value of the theoretical best, the
trace plot shows that it takes SA much longer to find the final basis.
This long interpolation path is partially due to the fluctuation in the
early iterations, where SA tends to generously accept inferior bases
before concentrating near the optimum. The PCA plot shows the
interpolation path and search points excluding the last termination
iteration. Pseudo-Derivative (PD) quickly gets stuck near the starting
position. Comparing the amount of random search done by CRS and SA, it
is surprising that SA does not carry as many samples as CRS. Combining
the insights from both the trace and PCA plot, one can learn the two
different search strategies by CRS and SA: CRS frequently samples in the
space and only make a move when an improvement is guaranteed to be made,
while SA first broadly accepts bases in the space and then starts the
extensive sampling in a narrowed subspace. The specific decision of
which optimiser to use will depend on the index curve in the basis space
but if the basis space is non-smooth, accepting inferior bases at first,
as SA has done here, can lead to a more efficient search, in terms of
the overall number of points evaluated.

@@

The next experiment compares the performance of CRS and SA when a local
maximum exists. Two search neighbourhood sizes: 0.5 and 0.7 are compared
to understand how a large search neighbourhood would affect the final
basis and the length of the search. Figure \ref{fig:kol-result} shows 80
paths simulated using 20 seeds in the PCA plot, faceted by the optimiser
and search size. With CRS and a search size of 0.5, despite being the
simplest and fastest, the optimiser fails in three instances where the
final basis lands neither near the local nor the global optimum. With a
larger search size of 0.7, more seeds have found the global maximum.
Comparing CRS and SA for a search size of 0.5, SA does not seem to
improve the final basis found, despite having longer interpolation
paths. However, the denser paths near the local maximum is an indicator
that SA is working hard to examine if there is any other optimum in the
basis space but the relatively small search size has diminished its
ability to reach the global maximum. With a larger search size, almost
all the seeds (16 out of 20) have found the global maximum and some
final bases are much closer to the theoretical best, as compared to the
three other cases. This indicates that SA, with a reasonable large
search window, is able to overcome the local optimum and optimise close
towards the global optimum.

@@

One interesting situation observed in the previous examples is that, for
some simulations, as shown on the left panel of Figure
\ref{fig:flip-sign}, the target basis is generated on the other half of
the basis space and the interpolator seems to draw a straight line to
interpolate. Bases with opposite signs do not affect the projection and
index value, but we would prefer the target to have the same orientation
as the current basis. The orientation of two bases can be
computationally checked by calculating the determinant -- a negative
value suggests the two bases having different orientation. For 1D bases,
this can be corrected by flipping the sign in one basis. For higher
dimensions, it can be a bit more difficult because the orthonormality of
the basis needs to be maintained also when an individual vector is
flipped. Here an orientation check is carried out once a new target
basis is generated and the sign in the target basis will be flipped if a
negative determinant is obtained. The interpolation after implementing
the orientation check is shown on the right panel of Figure
\ref{fig:flip-sign} where the unsatisfactory interpolation no longer
exists.

@@

\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis. Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation. The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
\end{figure}
\end{Schunk}

@@

  \begin{itemize}
  \tightlist
  \item
    \texttt{get\_anchor()} extracts target observations
  \item
    \texttt{get\_basis\_matrix()} flattens all the bases into a matrix
  \item
    \texttt{get\_best()} extracts the observation with the highest index
    value in the data object
  \item
    \texttt{get\_dir\_search()} extracts directional search observations
    for PD search
  \item
    \texttt{get\_interp()} extracts interpolated observations
  \item
    \texttt{get\_interp\_last()} extracts the ending interpolated
    observations in each iteration
  \item
    \texttt{get\_interrupt()} extracts the ending interpolated
    observations and the target observations if the interpolation is
    interrupted
  \item
    \texttt{get\_search()} extracts search observations
  \item
    \texttt{get\_search\_count()} extracts the count of search
    observations
  \item
    \texttt{get\_space\_param()} produces the coordinates of the centre
    and radius of the basis space
  \item
    \texttt{get\_start()} extracts the starting observation
  \item
    \texttt{get\_theo()} extracts the theoretical best observations, if
    given
  \end{itemize}
\item
  \texttt{bind\_*()} incorporates additional information outside the
  tour optimisation into the data object.

  \begin{itemize}
  \tightlist
  \item
    \texttt{bind\_theoretical()} binds the theoretical best observation
    in simulated experiment
  \item
    \texttt{bind\_random()} binds randomly generated bases in the
    projection bases space to the data object
  \item
    \texttt{bind\_random\_matrix()} binds randomly generated bases and
    outputs in a matrix format
  \end{itemize}
\item
  \texttt{add\_*()} provides wrapper functions to create ggprotos for
  different components for the PCA plot

  \begin{itemize}
  \tightlist
  \item
    \texttt{add\_anchor()} for plotting anchor bases
  \item
    \texttt{add\_anno()} for annotating the symmetry of start bases
  \item
    \texttt{add\_dir\_search()} for plotting the directional search
    bases with magnified distance
  \item
    \texttt{add\_end()} for plotting end bases
  \item
    \texttt{add\_interp()} for plotting the interpolation path
  \item
    \texttt{add\_interp\_last()} for plotting the last interpolation
    bases for comparing with target bases when interruption is used
  \item
    \texttt{add\_interrupt()} for linking the last interpolation bases
    with target ones when interruption is used
  \item
    \texttt{add\_search()} for plotting search bases
  \item
    \texttt{add\_space()} for plotting the circular space
  \item
    \texttt{add\_start()} for plotting start bases
  \item
    \texttt{add\_theo()} for plotting theoretical best bases, if
    applicable
  \end{itemize}
\item
  Utilities

  \begin{itemize}
  \tightlist
  \item
    \texttt{theme\_fern()} and \texttt{format\_label()} for better
    display of the grid lines and axis formatting
  \item
    \texttt{clean\_method()} to clean up the name of the optimisers
  \item
    \texttt{botanical\_palettes()} is a collection of colour palettes
    from Australian native plants. Quantitative palettes include daisy,
    banksia, and cherry and sequential palettes contain fern and acacia
  \item
    \texttt{botanical\_pal()} as the colour interpolator
  \item
    \texttt{scale\_color\_*()} and \texttt{scale\_fill\_*()} for scaling
    the colour and fill of the plot
  \end{itemize}
\end{itemize}

@@

This paper has provided several visual diagnostics that can be used for
understanding a complex optimisation procedure and are implemented in
the ferrn package. The methods were illustrated using the optimisers
available for projection pursuit guided tour. Here the constraint is the
orthonormality condition of the projection bases, which corresponds to
optimisation over spheres and torii. The approach described broadly
applies to other constrained optimisers. Although the manifold in
\(p\)-space might be different the diagnostic techniques are the same. A
researcher would begin by saving the path of the optimiser in a form
required to input into the ferrn package, as described in this paper.
One might generally make more samples from the constrained space upon
which to assess and compare the optimisation paths. These
high-dimensional data objects can then be viewed using the tour.

The progressive optimisation of a target function and it's coverage of
the search space can be viewed in both reduced 2D space and the full
space. These visualisations can lead to insights for evaluating and
comparing the performance of multiple optimisers operating on the same
task. They can provide better understanding of existing methods or
motivate the development of new approaches. For example, we have
compared how three optimisers perform when maximising a non-smooth index
function and illustrated how the pseudo-derivative search fails in this
setting. The observations from our experiments have also been translated
into improved optimisation methods for the guided tour, e.g.~we
introduced the option to interrupt the search if a better basis is found
along the path.

@@

This article is created using \CRANpkg{knitr} \citep{knitr} and
\CRANpkg{rmarkdown} \citep{rmarkdown} in R. The source code for
reproducing this paper can be found at:
\url{https://github.com/huizezhang-sherry/paper-tour-vis}.

>>>>

\caption[Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green)]{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green). The projection problem is a 2D projection with three variables using the holes index. The grey points are randomly generated 2D projection bases in the space, and it can be observed that these points form a torus. The full video of the animation can be found in the html version of the paper.}\label{fig:toy-torus}
\end{figure}
\end{Schunk}

@@

Random variables with different distributions have been simulated, and
the distributional form of each variable is presented in equations
\ref{eq:sim-norm} to \ref{eq:sim-x7}. Variables \texttt{x1}, \texttt{x8}
to \texttt{x10} are normal noise with zero mean, and unit variance and
\texttt{x2} to \texttt{x7} are normal mixtures with varied weights and
locations. All the variables have been scaled to have overall unit
variance before projection pursuit. The holes index
\citep{cook2008grand}, used for detecting bimodality of the variables,
is used throughout the examples unless otherwise specified.

@@

An example of non-monotonic interpolation has been given in Figure
\ref{fig:toy-interp}: a path that passes bases with a higher index value
than the target one. For SA, a non-monotonic interpolation is justified
since target bases do not necessarily have a higher index value than the
current one, while this is not the case for CRS. The original trace plot
for a 2D projection problem, optimised by CRS, is shown on the left
panel of Figure \ref{fig:interruption}, and one can observe that the
non-monotonic interpolation has undermined the optimiser to realise its
full potential. Hence, an interruption is implemented to stop at the
best basis found in the interpolation. The right panel of Figure
\ref{fig:interruption} shows the trace plot after implementing the
interruption, and while the first two interpolations are identical, the
basis at time 61 has a higher index value than the target in the third
interpolation. Rather than starting the next iteration from the target
basis on time 65, CRS starts the next iteration at time 61 on the right
panel and reaches a better final basis.

@@

\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS. On the left panel, the basis with a higher index value is found during the interpolation but not used. On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel.}\label{fig:interruption}
\end{figure}
\end{Schunk}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  the distance between the basis found and the current needs to be
  larger than 1e-3;
\item
  the relative change of the index value needs to be larger than 1e-5;
  and
\item
  the alpha parameter needs to be larger than 0.01.
\end{enumerate}

Figure \ref{fig:polish} presents the projected data and trace plot of a
2D projection, optimised by CRS and followed by the polish step. The top
row shows the initial projection, the final projection after CRS, and
the final projection after polish, respectively. The end basis found by
CRS reveals the four clusters in the data, but the edges of each cluster
are not clean-cut. Polish works with this end basis and further pushes
the index value to produce clearer edges of the cluster, especially
along the vertical axis.

@@

\caption{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index. The top row shows the initial projected data and the final views after CRS and polish search, and the second row traces the index value. The clustering structure in the data is detected by CRS (top middle panel), but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis. Note that the parameter \code{max.tries} is set to 400 in this experiment for CRS to do its best. }\label{fig:polish}
\end{figure}
\end{Schunk}

@@

The holes index function used for all the examples before this section
produces a smooth interpolation, while this is not the case for all the
indexes. An example of a noisy index function for 1D projections
compares the projected data, \(\mathbf{Y}_{n \times 1}\), to a randomly
generated normal distribution, \(\mathcal{N}_{n \times 1}\), using the
Kolmogorov test. Let \(F_{.}(n)\) be the empirical cumulative
distribution function (ECDF) with two possible subscripts, \(Y\) and
\(\mathcal{N}\), representing the projected and randomly generated data,
and \(n\) denoting the number of observations, the Kolmogorov index
\(I^{nk}(n)\), is defined as:

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  whether any optimiser fails to optimise this non-smooth index; and
\item
  whether the optimisers can find the global optimum despite the
  presence of a local optimum.
\end{enumerate}

Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of
all three optimisers, and as expected, none of the interpolated paths
are smooth. There is barely any improvement made by PD, indicating its
failure in optimising non-smooth index functions. While CRS and SA have
managed to get close to the index value of the theoretical best, the
trace plot shows that it takes SA much longer to find the final basis.
This long interpolation path is partially due to the fluctuation in the
early iterations, where SA tends to generously accept inferior bases
before concentrating near the optimum. The PCA plot shows the
interpolation path and search points, excluding the last termination
iteration. Pseudo-Derivative (PD) quickly gets stuck near the starting
position. Comparing the amount of random search done by CRS and SA, it
is surprising that SA does not carry as many samples as CRS. Combining
the insights from both the trace and PCA plot, one can learn the two
different search strategies by CRS and SA: CRS frequently samples in the
space and only make a move when an improvement is guaranteed to be made,
while SA first broadly accepts bases in the space and then starts the
extensive sampling in a narrowed subspace. The specific decision of
which optimiser to use will depend on the index curve in the basis
space, but if the basis space is non-smooth, accepting inferior bases at
first, as SA has done here, can lead to a more efficient search in terms
of the overall number of points evaluated.

@@

The next experiment compares the performance of CRS and SA when a local
maximum exists. Two search neighbourhood sizes, 0.5 and 0.7, are
compared to understand how a large search neighbourhood would affect the
final basis and the length of the search. Figure \ref{fig:kol-result}
shows 80 paths simulated using 20 seeds in the PCA plot, faceted by the
optimiser and search size. With CRS and a search size of 0.5, despite
being the simplest and fastest, the optimiser fails in three instances
where the final basis lands neither near the local nor the global
optimum. With a larger search size of 0.7, more seeds have found the
global maximum. Comparing CRS and SA for a search size of 0.5, SA does
not seem to improve the final basis found, despite having longer
interpolation paths. However, the denser paths near the local maximum
are an indicator that SA is working hard to examine if there is any
other optimum in the basis space, but the relatively small search size
has diminished its ability to reach the global maximum. With a larger
search size, almost all the seeds (16 out of 20) have found the global
maximum, and some final bases are much closer to the theoretical best,
as compared to the three other cases. This indicates that SA, with a
reasonable large search window, is able to overcome the local optimum
and optimise close towards the global optimum.

@@

One interesting situation observed in the previous examples is that, for
some simulations, as shown on the left panel of Figure
\ref{fig:flip-sign}, the target basis is generated on the other half of
the basis space, and the interpolator seems to draw a straight line to
interpolate. Bases with opposite signs do not affect the projection and
index value, but we would prefer the target to have the same orientation
as the current basis. The orientation of two bases can be
computationally checked by calculating the determinant -- a negative
value suggests the two bases have a different orientation. For 1D bases,
this can be corrected by flipping the sign on one basis. For higher
dimensions, it can be a bit more difficult because the orthonormality of
the basis needs to be also maintained when an individual vector is
flipped. Here, an orientation check is carried out once a new target
basis is generated, and the sign in the target basis will be flipped if
a negative determinant is obtained. The interpolation after implementing
the orientation check is shown on the right panel of Figure
\ref{fig:flip-sign}, where the unsatisfactory interpolation no longer
exists.

@@

\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis. Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463. The dots represent the target basis in each iteration, and the path shows the interpolation. On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space), and the interpolator crosses the origin to perform the interpolation. The right panel shows the same interpolation after implementing an orientation check, and the undesirable interpolation disappears.}\label{fig:flip-sign}
\end{figure}
\end{Schunk}

@@

  \begin{itemize}
  \tightlist
  \item
    \texttt{get\_anchor()} extracts target observations.
  \item
    \texttt{get\_basis\_matrix()} flattens all the bases into a matrix.
  \item
    \texttt{get\_best()} extracts the observation with the highest index
    value in the data object.
  \item
    \texttt{get\_dir\_search()} extracts directional search observations
    for PD search.
  \item
    \texttt{get\_interp()} extracts interpolated observations.
  \item
    \texttt{get\_interp\_last()} extracts the ending interpolated
    observations in each iteration.
  \item
    \texttt{get\_interrupt()} extracts the ending interpolated
    observations and the target observations if the interpolation is
    .interrupted
  \item
    \texttt{get\_search()} extracts search observations.
  \item
    \texttt{get\_search\_count()} extracts the count of search
    observations.
  \item
    \texttt{get\_space\_param()} produces the coordinates of the centre
    and radius of the basis space.
  \item
    \texttt{get\_start()} extracts the starting observation.
  \item
    \texttt{get\_theo()} extracts the theoretical best observations, if
    given.
  \end{itemize}
\item
  \texttt{bind\_*()} incorporates additional information outside the
  tour optimisation into the data object.

  \begin{itemize}
  \tightlist
  \item
    \texttt{bind\_theoretical()} binds the theoretical best observation
    in simulated experiment.
  \item
    \texttt{bind\_random()} binds randomly generated bases in the
    projection bases space to the data object.
  \item
    \texttt{bind\_random\_matrix()} binds randomly generated bases and
    outputs in a matrix format.
  \end{itemize}
\item
  \texttt{add\_*()} provides wrapper functions to create ggprotos for
  different components for the PCA plot

  \begin{itemize}
  \tightlist
  \item
    \texttt{add\_anchor()} for plotting anchor bases.
  \item
    \texttt{add\_anno()} for annotating the symmetry of start bases.
  \item
    \texttt{add\_dir\_search()} for plotting the directional search
    bases with magnified distance.
  \item
    \texttt{add\_end()} for plotting end bases.
  \item
    \texttt{add\_interp()} for plotting the interpolation path.
  \item
    \texttt{add\_interp\_last()} for plotting the last interpolation
    bases for comparing with target bases when interruption is used.
  \item
    \texttt{add\_interrupt()} for linking the last interpolation bases
    with target ones when interruption is used.
  \item
    \texttt{add\_search()} for plotting search bases.
  \item
    \texttt{add\_space()} for plotting the circular space.
  \item
    \texttt{add\_start()} for plotting start bases.
  \item
    \texttt{add\_theo()} for plotting theoretical best bases, if
    applicable.
  \end{itemize}
\item
  Utilities

  \begin{itemize}
  \tightlist
  \item
    \texttt{theme\_fern()} and \texttt{format\_label()} for better
    display of the grid lines and axis formatting.
  \item
    \texttt{clean\_method()} to clean up the name of the optimisers.
  \item
    \texttt{botanical\_palettes()} is a collection of colour palettes
    from Australian native plants. Quantitative palettes include daisy,
    banksia, and cherry, and sequential palettes contain fern and
    acacia.
  \item
    \texttt{botanical\_pal()} as the colour interpolator.
  \item
    \texttt{scale\_color\_*()} and \texttt{scale\_fill\_*()} for scaling
    the colour and fill of the plot.
  \end{itemize}
\end{itemize}

@@

This paper has provided several visual diagnostics that can be used for
understanding a complex optimisation procedure and are implemented in
the \textbf{ferrn} package. The methods were illustrated using the
optimisers available for projection pursuit guided tour. Here the
constraint is the orthonormality condition of the projection bases,
which corresponds to optimisation over spheres and torii. The approach
described broadly applies to other constrained optimisers. Although the
manifold in \(p\)-space might be different the diagnostic techniques are
the same. A researcher would begin by saving the path of the optimiser
in a form required to input into the ferrn package, as described in this
paper. One might generally make more samples from the constrained space
upon which to assess and compare the optimisation paths. These
high-dimensional data objects can then be viewed using the tour.

The progressive optimisation of a target function and its coverage of
the search space can be viewed in both reduced 2D space and the full
space. These visualisations can lead to insights for evaluating and
comparing the performance of multiple optimisers operating on the same
task. They can provide a better understanding of existing methods or
motivate the development of new approaches. For example, we have
compared how three optimisers perform when maximising a non-smooth index
function and have illustrated how the pseudo-derivative search fails in
this setting. The observations from our experiments have also been
translated into improved optimisation methods for the guided tour, e.g.,
we introduced the option to interrupt the search if a better basis is
found along the path.

@@

This article is created using \CRANpkg{knitr} \citep{knitr} and
\CRANpkg{rmarkdown} \citep{rmarkdown} in R. The source code for
reproducing this paper can be found at:
\url{https://github.com/huizezhang-sherry/paper-ferrn}.
